%
% Documento: Abstract
%

\begin{resumo}[Abstract]

This work develops and evaluates a passive UHF RFID reader prototype for asset inventory, comprising a YPD-R200 module, a linearly polarized 5~dBi Airplux APCA8090 antenna, AD-239 M730 tags, and an Arduino Uno. The proposal replaces an initial NFC-based design whose short range requires each item to be checked individually. The reader library was adapted to the Arduino TTL UART through the \texttt{Stream} abstraction and \texttt{SoftwareSerial}, enabling variable-length EPC acquisition and accumulation of unique identifiers within each observation window. The main campaign was conducted on July 13, 2026, in an enclosed institutional laboratory of approximately 50~m$^2$ and included distance, orientation, mounting-material, and inventory tests. Across 15 attempts per distance, detection rates were 100.0\% at 0.5~m, 73.3\% at 1.0~m, 66.7\% at 1.5~m, and 53.3\% at 2.0~m. At 1.5~m, rates were 93.3\%, 86.7\%, and 66.7\% at $0^\circ$, $45^\circ$, and $90^\circ$, respectively. A complementary isolated-tag test with 30 attempts per combination yielded 100.0\%, 86.7\%, and 60.0\% at 1.0~m for the same angles; 90.0\%, 70.0\%, and 36.7\% at 2.0~m; and 0.0\% at 5.0~m for all three orientations. These results show that some failures occur even without simultaneous responses from multiple tags. Mounting tests yielded 100.0\% on cardboard, 0.0\% on direct metal, and 100.0\% on metal with a 10~mm plastic spacer. Across five inventory rounds with five tags, mean coverage was 76.0\%, but no round identified every asset. The observed complete-round rate was 0.0\%, with a 95\% Wilson confidence interval from 0.0\% to 43.4\%. Because gradual expansion to approximately 50 tags depended on complete coverage at the initial stage, scaling was not performed. Budget constraints motivated the selected hardware, but the results show that the current configuration should not be adopted as an operational reference for one-pass inventory. It remains useful as an experimental baseline for improvements. Future work should compare circularly polarized or polarization-diverse antennas, assess the azimuth coverage of omnidirectional alternatives, and evaluate on-metal tags before repeating the scaling procedure.

\textbf{Keywords:} UHF RFID; asset inventory; YPD-R200; experimental measurements; RFID antennas.

\end{resumo}
